---
title: 'IS 606: Assignment 8'
author: "MUSA T. GANIYU"
date: "March 17, 2016"
output: pdf_document
---
**QUESTION:**

**Pick three of your favorite books on one of your favorite subjects. At least one of the books should have more**
**than one author. For each book, include the title, authors, and two or** **three other attributes that you find**
**interesting.**
**Take the information that you’ve selected about these three books, and separately create three files which**
**store the book’s information in HTML (using an html table), XML, and JSON formats (e.g. “books.html”,**
**“books.xml”, and “books.json”). To help you better understand the different file structures, I’d prefer that you**
**create each of these files “by hand” unless you’re already very comfortable with the file formats.**

**DIRECTION:**

**Write R code, using your packages of choice, to load the information from each of the three sources into separate R data frames.**


**XML Format**

```{r}
# Please kindly install these packages if you dont have it yet, and load it.
library(XML)
library(RCurl)
library(knitr)
library(htmltab)
library(jsonlite)

# Getting xml data loaded from URL

url = "https://raw.githubusercontent.com/mascotinme/MSDA-IS607/f896255f3b817cad1496bb767922bb7c72457422/book.xml"
data = getURL(url, ssl.verifypeer=0L, followlocation=1L)
book1 <- xmlParse(data)
book <- xmlToDataFrame(book1)
kable(head(book))

```

**HTML FORMAT**

```{r}
# load HTML data from URL and ingnore the warning statement(s).
options(warn = -1)

url = "https://raw.githubusercontent.com/mascotinme/MSDA-IS607/master/Books1.html"

books_html = htmltab(doc = url, which = 1)
books_html = data.frame(books_html, row.names = NULL)
kable(books_html)



```

**JSON FORMAT**

```{r}
#loading json file from URL

url2 = "https://raw.githubusercontent.com/mascotinme/MSDA-IS607/master/Books.json"

books_json = getURL(url2)
books_json = fromJSON(books_json)
kable(books_json)
```

**Question: Are the three data frames identical?**

**YES!, the three data frames are identical.**





**REFERENCES:**

**http://www.rprimer.dk/index.php?page=showrule&index=2**

**https://cran.r-project.org/web/packages/jsonlite/vignettes/json-aaquickstart.html**

**http://www.w3schools.com/xml/**
