---
title: 'IS 607: Assignment 10'
author: "MUSA T. GANIYU"
date: "March 29, 2016"
output:
  pdf_document:
    toc: yes
  html_document:
    code_folding: hide
    highlight: pygments
    theme: cerulean
    toc: yes
    toc_depth: 4
    toc_float:
      collapsed: yes
      smooth_scroll: no
---
Your task is to choose one of the New York Times APIs, construct an interface in R to read in the JSON d
ata, and 
transform it to an R dataframe

Note: Load the require packages for easy accesibility.

```{r}
library(knitr)
library(XML)
library(jsonlite)
library(plyr)

```

Load the data file from the New York Times webpage. And an API key is required for accessing the webpage as a developer.

```{r}
url.times <- ("http://api.nytimes.com/svc/politics/v3/us/legislative/congress/107-113/nominees/updated.xml?api-key=685874ff71bb286631a8ea2c3f9989bb:16:74820193")

XML_doc   <- htmlParse(url.times)

str(XML_doc)
```

We found out that the dataset isnt in dataframe yet, we will therefore convert if from list to dataframe.


```{r}


nytimes_all <- ldply(xmlToList(XML_doc), data.frame)
```

As you have noticed, we have a whole bunch of "repeating" columns, we will therefore subset 10 of it and rename it.

```{r}
nytimes <- nytimes_all[, 1:10]

names(nytimes)[names(nytimes)==".id"] <- "id"
names(nytimes)[names(nytimes)=="result_set.status"] <- "status"
names(nytimes)[names(nytimes)=="result_set.copyright"] <- "copyright"
names(nytimes)[names(nytimes)=="result_set.results.congress"] <- "congress_results"
names(nytimes)[names(nytimes)=="result_set.results.num_results"] <- "num_results"
names(nytimes)[names(nytimes)=="result_set.results.nominations.nomination.id"] <- "nomination_id"
names(nytimes)[names(nytimes)=="result_set.results.nominations.nomination.uri"] <- "nomination_uri"
names(nytimes)[names(nytimes)=="result_set.results.nominations.nomination.date_received"] <- "date_received"
names(nytimes)[names(nytimes)=="result_set.results.nominations.nomination.description"] <- "description"
names(nytimes)[names(nytimes)=="result_set.results.nominations.nomination.nominee_state"] <- "state"

xmlSize(nytimes) #how many children in node, 10
nytimes[[1]]
```

Here is the extracted dataset head


```{r}

names(nytimes)
kable(head(nytimes))
```


lets now try to access thesame dataset from different format called (JSON)

```{r}

url.json <- ("http://api.nytimes.com/svc/politics/v3/us/legislative/congress/107-113/nominees/updated.json?api-key=685874ff71bb286631a8ea2c3f9989bb:16:74820193")

# We have to extract the file from url to json format using fromJson

nytimes.json <- fromJSON(url.json)

# Setting it to dataframe for analysis

nytimes.json <- ldply (nytimes.json[4], data.frame)

kable(head(nytimes.json))


```

