---
title: 'IS 607: PROJECT'
author: "MUSA T. GANIYU"
date: "April 28, 2016"
output:
  html_document:
    highlight: pygments
    theme: cerulean
  pdf_document: default
  tidy: yes
---


> **ABSTRACT:**

* The project analysis is done on three (3) different datasets from different sources. The first dataset if from the University of California, Irvine website on **Air Quality**, the second dataset was extracted as an .xls file from World Bank website for the year **1961-1999**, while the third dataset was base on incident rate of leukemia in US States and was extracted from cancer.gov website.

* Analyses was done on all the three datasets to determine if the is linear relationship between them. I found out that the Air Quality has some of it variables that are way beyond the acceptable threnshold, I checked for interaction between those variables and benzene. I compared the result with the Leukemia incident rate in the 49 US States and the result is amazing. 

* We would **randomly** select observations to avoid bias.

* We would do some of the analysis on three(3) different programs; R, MySQL and MongoDB.


> **AIM AND OBJECTIVE:**

  * I am motivated by the serious impacts of the Air Quality throughout the whole world. Lots of adults, children are prone or exposed to the various categories of air throughout the globe. I thought it would be good to do an analysis to see the level of impact and to offer solutions.


> Load Libraries:

* Please, kindly install the package(s) first if you don't already have it and load their respective libraries.

```{r}

options(warn = -1)
suppressMessages(library(xlsx))
library(xlsxjars)
library(rJava)
library(devtools)
suppressMessages(library(RMySQL))
library(DBI)
suppressMessages(library(plotly))
suppressWarnings(library(rmongodb))
library(knitr)
library(mongolite)
suppressMessages(library(RCurl))
suppressMessages(library(plyr))
suppressMessages(library(ggplot2))
```


> We will be using the Air Quality dataset from the University of California, Irvine website: 

http://archive.ics.uci.edu/ml/datasets/Air+Quality#

** It is a .xlsx dataset that I downloaded into my local computer.**

+ **Kindly download from here:**

https://github.com/mascotinme/MSDA-IS607/blob/master/cckp_historical_data_0.xls

```{R}
airqu <- read.xlsx2("C:/Users/mayowa/Documents/GitHub/MSDA-IS607/AirQualityUCI.xlsx", sheetName = "AirQualityUCI", header=TRUE, colClasses=NA)

dim(airqu)
kable(head(airqu))
kable(tail(airqu))
names(airqu)
str(airqu)
class(airqu)
```


> * We notice that there are alot of NA's in the dataset, we therefore thought it would be appropriate (for easy analysis) to replace the missing values or NA with Zero(o)
  
  * We also notice that the column names is ambiguous, we therefore change it to a reasonable column names.

```{R}

airqu$NMHC.GT.[is.na(airqu$NMHC.GT.)] <- 0.0000
airqu$CO.GT.[is.na(airqu$CO.GT.)] <- 0.0000
airqu$PT08.S1.CO.[is.na(airqu$PT08.S1.CO.)] <- 0.0000
airqu$C6H6.GT.[is.na(airqu$C6H6.GT.)] <- 0.0000
airqu$PT08.S2.NMHC.[is.na(airqu$PT08.S2.NMHC.)] <- 0.0000
airqu$NOx.GT.[is.na(airqu$NOx.GT.)] <- 0.0000
airqu$PT08.S3.NOx.[is.na(airqu$PT08.S3.NOx.)] <- 0.0000
airqu$NO2.GT.[is.na(airqu$NO2.GT.)] <- 0.0000
airqu$PT08.S4.NO2.[is.na(airqu$PT08.S4.NO2.)] <- 0.0000
airqu$PT08.S5.O3.[is.na(airqu$PT08.S5.O3.)] <- 0.0000
airqu$T[is.na(airqu$T)] <- 0.0000
airqu$RH[is.na(airqu$RH)] <- 0.0000
airqu$AH[is.na(airqu$AH)] <- 0.0000

kable(head(airqu))


colnames(airqu)[1] <- "date"
colnames(airqu)[2] <- "time"
colnames(airqu)[3] <- "co"
colnames(airqu)[4] <- "tin_oxide"
colnames(airqu)[5] <- "hydrocarbon"
colnames(airqu)[6] <- "benzene"
colnames(airqu)[7] <- "titania"
colnames(airqu)[8] <- "nitroge_oxide"
colnames(airqu)[9] <- "tungsten-oxide"
colnames(airqu)[10] <- "no2"
colnames(airqu)[11] <- "tungsten_oxide_o2"
colnames(airqu)[12] <- "indium_oxide"
colnames(airqu)[13] <- "temp_ac"
colnames(airqu)[14] <- "rh"
colnames(airqu)[15] <- "ah"


kable(head(airqu))
````



> We will now connect R with MySql database.


```{R}

mysql = dbConnect(MySQL(), user='root', password='oracle', dbname='world', host='localhost') # This connects the R with MySQL. Note that the login entities may be different from your login details, kindly adjust as neccesary.
```


>We would like to view the list of database(s) which are already in MySQL database.


```{R}
class(mysql)
dbListTables(mysql)
```

>Dropping and adding of table if they already exist to give access to the new table to be created. 

```{R}

dbSendQuery(mysql, 'drop table if exists airqu')



dbWriteTable(conn=mysql, name='airqu', value=airqu)
dbListFields(mysql, 'airqualityuci')

```


>We can now read and access the newly created table in R from MySQL directly.

```{R}

airqualityuci <- dbReadTable(conn=mysql, name='airqu')

write.table(airqualityuci, file = "airqualityuci.csv",row.names=FALSE, na="",col.names=TRUE, sep=",")

```


> **Attribute Information :**

+  **Date:** Date (DD/MM/YYYY)                     
+ **Time:** Time (HH.MM.SS)            

+ **co:** True hourly averaged concentration CO in mg/m^3

+ **tin_oxide:** PT08.S1 (tin oxide) hourly averaged sensor response (nominally CO targeted)   

+ **hydrocarbon:** True hourly averaged overall Non Metanic HydroCarbons concentration in microg/m^3 (reference analyzer) 

+ **benzene:** True hourly averaged Benzene concentration in microg/m^3 (reference analyzer)

+ **titania:**  PT08.S2 (titania) hourly averaged sensor response (nominally NMHC targeted)

+ **nitroge_oxide:** True hourly averaged NOx concentration in ppb (reference analyzer) 

+ **tungsten_oxide:** PT08.S3 (tungsten oxide) hourly averaged sensor response (nominally NOx targeted) 

+ **no2:** True hourly averaged NO2 concentration in microg/m^3 (reference analyzer) 

+ **tungsten_oxide2:** PT08.S4 (tungsten oxide) hourly averaged sensor response (nominally NO2 targeted)

+ **indium_oxide:** PT08.S5 (indium oxide) hourly averaged sensor response (nominally O3 targeted)

+ **temp_ac:** Temperature in °C

+ **rh:** Relative Humidity (%)

+ **ah:** AH Absolute Humidit

** we are going to randomly sample 900 (Less than 10%) observations from the total dataset.**

```{R}

airqualityuci <- airqualityuci[sample(nrow(airqualityuci), 900), ]
dim(airqualityuci)
str(airqualityuci)
summary(airqualityuci)


```

> **What recommendations has the federal government made to protect human health?**


....EPA estimates that 10 ppb benzene in drinking water that is consumed regularly or exposure to 0.4 ppb in air over a lifetime could cause a risk of one additional cancer case for every 100,000 exposed persons......


+ The underlisted output displays the date, time of benzene with ppb greater than the safely regulated ppb for benzene.

```{R}


mysqlquery <- dbGetQuery(mysql, "SELECT Date, Time, benzene 
     FROM airqu 
     WHERE benzene IN (SELECT benzene 
                  FROM airqu 
                  WHERE benzene > 10)")

kable(head(mysqlquery))

```

* No wonder! The EPA has classified benzene as a Group A, known human carcinogen!

+ https://www3.epa.gov/airtoxics/hlthef/benzene.html


> **MONGODB ---A NO-SQL**

* use this query from your cmd shell: 

mongoimport --db world --collection airqualityuci --type csv --file"c:\Users\mayowa\Documents\airqualityuci.csv" --headerline

* Adjust the file location accordingly

```{R}
mongo_conn <- mongo.create() # createing a MongoDB connection

mongo.is.connected(mongo_conn)  # checking if the mongo is connected


if(mongo.is.connected(mongo_conn) == TRUE) {
  mongo.get.databases(mongo_conn)
}


if(mongo.is.connected(mongo_conn) == TRUE) {
  db <- "world"
  mongo.get.database.collections(mongo_conn, db)
}


mongo_conn = mongo(collection = "airqualityuci",  db = "world")
mongo_conn$insert(airqualityuci) # Insert the airqualityuci into mongodb collection

```



```{r}

plot(airqualityuci$temp_ac, airqualityuci$benzene, main = "Scatter Plot", xlab = "Temperature", ylab = "Benzene")
abline(lm(airqualityuci$temp_ac~airqualityuci$benzene), col = "blue");

cor(airqualityuci$temp_ac, airqualityuci$benzene)


temp_co <- lm(temp_ac ~ benzene, data=airqualityuci)

summary.lm(temp_co)


plot(x=temp_co$residuals, y=temp_co$benzene)
abline(h = 0, lty = 3, col = "blue")

hist(temp_co$residuals)



airqual <- subset(airqualityuci, select=c(co, tin_oxide, hydrocarbon, benzene, titania, nitroge_oxide, tungsten.oxide, no2, tungsten_oxide_o2, indium_oxide, temp_ac, rh, ah  ))

plot(airqual)
```


> We are going to do  a Multiple Linear Regression model. But, we will want to choose the best model for our prediction.


${ Y }\quad =\quad { B }_{ 0 }\quad +\quad { B }_{ 1 }{ X }_{ 1 }\quad +\quad { B }_{ 2 }{ X }_{ 2 }\quad +\quad$ .........+$\quad { B }_{ n }{ X }_{ n }\quad$ +$\quad { e}\\$




> **SETTING UP HYPOTHESIS: # 1**

* Null Hypothesis: The population parameters(means) are the same

    $H_o:$ $\mu_1$ = $\mu_2$ = $\mu_3$.....= $\mu_n$ 


* Alternative Hythesis: Not $H_o$
   
   $H_a:$  Not $H_o$  : $\mu_1$ $\neq$ $\mu_2$ $\neq$  $\mu_3$.....$\neq$ $\mu_n$


>  **Rejection**:

Reject $H_o$ (Null Hypothesis) if the calculated value (P-Value) is less that the tabulated value(Table value = *0.05* ), otherwise do not reject $H_o$



* Underlisted are the best variables for the model, and analysis would be made on them.


```{r}
modelselection <- lm(temp_ac ~ tin_oxide+hydrocarbon+benzene+titania+nitroge_oxide+tungsten.oxide+no2+tungsten_oxide_o2+indium_oxide+co+rh+ah, data=airqual)

stepwise <- step(modelselection, direction = "both") # Model Selection using both FORWARD AND BACKWARD selection.

summary(stepwise)

```


* We will like to further test by using another statistical tool called Analysis of Variance (ANOVA)

> **ANALYSIS OF VARIANCE (ANOVA)**

```{r}
anova(stepwise, test= "F")
confint(stepwise)

```


> The Error Term or Residual

* We are going to check the distribution of the residuals with normal dataset distribution to determine the stochastics difference between the expected value and the observed value.

```{R}

residual_temp = data.frame(Fitted = fitted(stepwise),
Residuals = resid(stepwise), Treatment = airqual$temp_ac)

residual_benzene = data.frame(Fitted = fitted(stepwise),
Residuals = resid(stepwise), Treatment = airqual$benzene)



ggplot(residual_temp, aes(Fitted, Residuals, colour = Treatment)) + geom_point()
ggplot(residual_benzene, aes(Fitted, Residuals, colour = Treatment)) + geom_point()


```


**Decision & Conclusion:**

From the above multiple regression analysis and ANOVA, we found that the P-value is actually low compare to the Table value (0.05), we therefor reject $H_o$ (Null hypothesis) and conclude that there is/are difference(s) between the variables.




> **SECOND DATASET WAS EXTRACTED FROM THE WORLD BANK WEBSITE AS A .XLS FILE** 
 
 * This dataset contains historical temperature extracted from the Climate Research Unit (Mitchell et al, 2003), aggregated to the country and basin levels.  The dataset time is from **1961-1999**.  
 
 * Country_temperatureCRU:  mean monthly and annual temperatures by country for the period 1961-1999.  Values are in degrees Celsius.

 * http://databank.worldbank.org/data/download/catalog/cckp_historical_data_0.xls

+ we will like to compare the dataset with the above airquality analysis.

```{R}
suppressMessages(library(dplyr))

dataset2 <- read.xlsx("C:/Users/mayowa/Documents/cckp_historical_data_0.xls", sheetIndex=4,
  startRow=NULL, endRow=179, colIndex=NULL,
  as.data.frame=TRUE, header=TRUE,
  keepFormulas=FALSE)

str(dataset2)

kable(summary(dataset2))

```


> we are going to sample 178 observations from the 1st dataset to match with the second dataset to avoid bias.


**SETTING UP HYPOTHESIS: # 2**

* Null Hypothesis: There is no difference between means temperature in dataset 1 and dataset 2

    $H_o:$ $\mu_1$ = $\mu_2$ = $\mu_3$.....= $\mu_n$ 


* Alternative Hythesis:
   
    $H_a:$  Not $H_o:$ $\mu_1$ $\neq$ $\mu_2$ $\neq$  $\mu_3$.....$\neq$ $\mu_n$

**Rejection**:

Reject $H_o$ (Null Hypothesis) if the calculated value (P-Value) is less that the tabulated value(Table value = *0.05* ), otherwise do not reject $H_o$


**Decision:**

We therefore Accept $H_o$, since calculated calculated value (P-Value) is greater than the tabulated value(Table value = *0.05* )

```{R}
sample1 <- airqual[sample(nrow(airqual), 178), ]
dim(sample1)


temp_2 <- lm(sample1$temp_ac ~ dataset2$Annual_temp)

summary.lm(temp_2)

cor.test(sample1$temp_ac, dataset2$Annual_temp, method = c("pearson", "kendall", "spearman"),
         exact = NULL, conf.level = 0.95, continuity = FALSE)


anova(temp_2)

```


> **DATASET THREE(3)**

* The dataset is a Leukemia incident rate collected from the Center for Diseases Control and Prevention Website for 49 US States.....

** Data Source:**

* http://www.statecancerprofiles.cancer.gov/incidencerates/index.php?stateFIPS=00&cancer=090&race=00&sex=0&age=009&year=0&type=incd&sortVariableName=rate&sortOrder=default#results


**SETTING UP HYPOTHESIS: # 3**

* Null Hypothesis: There is no relationship between benzene in dataset 1 and and the average incident in dataset 3

    $H_o:$ $\mu_1$ = $\mu_2$ = $\mu_3$.....= $\mu_n$ 


* Alternative Hythesis:
   
    $H_a:$  Not $H_o:$ $\mu_1$ $\neq$ $\mu_2$ $\neq$  $\mu_3$.....$\neq$ $\mu_n$

**Rejection**:

Reject $H_o$ (Null Hypothesis) if the calculated value (P-Value) is less than the tabulated value(Table value = *0.05* ), otherwise do not reject $H_o$


```{R}
library(plyr)

incd <- read.csv("https://raw.githubusercontent.com/mascotinme/MSDA-IS607/master/incd.csv", sep = ",", header = TRUE, skip = 8)
 
str(incd)
# Rename the column names for easy accesibility

incd$Average.Annual.Count <- as.numeric(incd$Average.Annual.Count)

names(incd) <- c("state", "FIPS", "adj_age_100000", "lower_Adj_CI", "upper_Adj_CI", "avg_annual_count", "Recent.Trend", "5years_rate", "Lower_5Yrs_CI", "upper_5Yrs_CI")



# We are going to seperate the rows(Stable, Rising and Falling)


str(incd)

kable(head(incd))

stable <- subset(incd, incd$Recent.Trend == "stable")
rising <- subset(incd, incd$Recent.Trend == "rising")
falling <- subset(incd, incd$Recent.Trend == "falling")

kable(head(stable))
kable(head(rising))
kable(head(falling))


require(ggplot2)

ggplot(stable, col = "red", aes(x =state, y=round(avg_annual_count,2))) +
       geom_bar(stat="identity",position="dodge") + 
       xlab("States") + 
       ylab("Average Annual counts") + 
       ggtitle("Average Annual Stable Leukemia Counts per US States")


ggplot(rising, col = "blue", aes(x =state, y=round(avg_annual_count,2))) +
       geom_bar(stat="identity",position="dodge") + 
       xlab("States") + 
       ylab("Average Annual counts") + 
       ggtitle("Average Annual Rising Leukemia Counts per US States")


sample_all1 <- sample1[sample(nrow(sample1), 45), ]
sample_all2 <- dataset2[sample(nrow(dataset2), 45), ]
sample_all3 <- incd[sample(nrow(incd), 45), ]
allsamples <- data.frame(c(sample_all1, sample_all2, sample_all3))


all_dataset_subset <- allsamples[, c("avg_annual_count", "Annual_temp" , "co", "benzene")]
str(all_dataset_subset)




linearmodel <- lm(all_dataset_subset$avg_annual_count ~(all_dataset_subset$co + all_dataset_subset$benzene))
linearmodel

anova(linearmodel)
```


**Decision:**

We therefore Reject $H_o$, since calculated calculated value (P-Value) is greater than the tabulated value(Table value = *0.05* )



> **GENERAL CONCLUSION:**


Based on the above analyzes done on the three datasets, we therefore conclude that there are strong relationships between the carbon-monoxide (co), Temperature (temp_ac) and Benzene. Though, the datasets might not be enough to be able to determine the actual causes,  and links between the variables, we are strongly confident that a reduction in the exposure to their atmospheric condition (particularly Benzene) may aid in reducing leukemia in the United States and the World at large.

Note: The analysis is not exclusive as some other contributing factor(s) may also needs to be accessed.

```{}

```



> REFERENCES:

* http://star.psy.ohio-state.edu/coglab/Demos/r.anova.html

* http://archive.ics.uci.edu/ml/datasets/Air+Quality# (Data Source)

* http://www.statecancerprofiles.cancer.gov/incidencerates/index.php?stateFIPS=00&cancer=090&race=00&sex=0&age=009&year=0&type=incd&sortVariableName=rate&sortOrder=default#results (Data Source)


* https://www.youtube.com/watch?v=1NOlgZLoojE 

* http://stackoverflow.com/questions/7531868/how-to-rename-a-single-column-in-a-data-frame-in-r

* http://databank.worldbank.org/data/download/catalog/cckp_historical_data_0.xls (Data source)

* http://www.ats.ucla.edu/stat/r/faq/subset_R.htm