---
title: 'IS 607: WEEK 4 ASSIGNMENT SOLUTION'
author: "MUSA T. GANIYU"
date: "February 20, 2016"
output: pdf_document
---
3. We load the data from example given in chapter 8 of Automated Data Collection with R (page 196).

```{r}
data <- "555-123Moe Szyslak (636) 555-0113Burns, C. Montgomery555-6542Rev. Timothy Lovejoy555 8904Ned Flanders636-555-3226Simpson, Homer555364Dr. Julius Hibbert";

library(stringr);

name <- unlist(str_extract_all(data, "[[:alpha:]., ]{2,}"))

name;
 # Rearrange the vector to so that all element conform to the standard first_name, last_name.

sort(name, partial = NULL, na.last = NA, decreasing = FALSE,
     method = c("first_name", "last_name"), index.return = FALSE);

# Vector indicating wether a character has a title ( i.e Rev. and Dr.)

str_extract(name, ("Dr.|Rev."));
str_detect(name, ("Dr.|Rev."));

# Vector indicating wether a character has a second name.

str_detect(name, ("second name"));

```

4.
Consider the string < title>+++BREAKING NEWS+++</title>. We would like to extract the first HTML tag. To do so we write the regular expression <.+>. Explain why this fail and correct the expression.    

```{r}
# note that this is HTML with + as COMMON QUANTIFICATION OPERATOR, "." as character to extract order in sequence.



html_tag <- "< title>+++BREAKING NEWS+++</title>";
str_extract(html_tag, "<.+>");


# This is a Greedy Quantification; We Correct this by adding the operator "?" after operator "+".

str_extract(html_tag, "<.+?>");

```
8. Consider the string (5-3)^2=5^2-2*5*3+3 conforms to the binomial theorem. We would like to extract the formula in the string. To do so we write the regular expression [^0-9=+*()] +.Explain why this fails and correct the expression.


```{r}
data2 <- "(5-3)^2=5^2-2*5*3+3 conforms to the binomial theorem.";

str_extract(data2, "[^0-9=+*()]+");

# The "^" raises all the characters at its end, and the "-" makes an inclusion in the character class.

str_extract(data2, "[0-9=+*()^]+");

str_extract(data2, "[0-9=+*()^-]+")

```

